\section{Finance formulas  -- Practice exercises}

\bigskip

Formulas referenced in the worksheets:
 \bigskip
 
 \bigskip
 \framebox{
 \begin{minipage}[c]{.85\textwidth}  
~ \bigskip \\  \textsc{Compound Interest Formula:} \quad
$\displaystyle a = p \left( 1 + \frac{r}{12}\right) ^{12y}$ 

~ \bigskip \\  \textsc{Equivalent APR Formula:} \quad 
$\displaystyle \text{APR} = \left(1+\frac{r}{12}\right)^{12}-1$ 
 \bigskip

~ \bigskip \\  \textsc{Future Value Annuity Formula:} \quad
$\displaystyle a = p \ast \frac{\left( 1 + \frac{r}{12}\right) ^{12y}-1}{\frac{r}{12}}$ 
\bigskip

~ \bigskip \\  \textsc{Loan Payment Formula:} \quad
$\displaystyle p = \frac{a  \ast \frac{r}{12}}{1-\left( 1 + \frac{r}{12}\right) ^{-12y}}$ 

 \bigskip

~\quad where 
\begin{center}
\begin{tabular} {l} 

$a$ = account balance or loan amount (\$) \\
$p$ = initial deposit (principal), regular deposit, or regular payment (\$) \\
$y$ = time invested (years)\\
$r$ = interest rate compounded monthly (as a decimal) \\ 
\end{tabular}
\end{center}
\bigskip

\end{minipage}
}
\bigskip

\newpage %%%%%%

\begin{enumerate}
\item Use the indicated formulas to help Kiran figure out her finances.
%\item (a) \$3,305.13 \quad (b) 7.23\% \quad (c) \$415,475 \quad (d) \$400.65
\begin{enumerate}
\item Kiran deposited \$\text{2,500} in a money market account that earned 7\% interest compounded monthly.  Use the \textsc{Compound Interest Formula} to calculate her account balance after 4 years. \vfill
\item What is the equivalent APR on Kiran's money market account?  Use the \textsc{Equivalent APR Formula.} \vfill
\item Kiran puts \$400 a month in her retirement account that amazingly also earns 7\% interest compounded monthly.  Use the \textsc{Future Value Annuity Formula} to determine how much Kiran will have in her retirement account in 28 years. \vfill
\item Kiran would really like to buy a new hybrid car that sells for \$\text{23,500}.  Sadly Kiran's credit rating is not very good, so the best the dealership offers is a loan at (you guessed it) 7\% interest compouned monthly.  Use the \textsc{Loan Payment Formula} to calculate her monthly car payments on a six year loan. \vfill
\end{enumerate}

\newpage %%%%%%

\item Tim and Josh are saving for their kids' college in fifteen years. The account pays the equivalent of 5.4\% interest compounded monthly (taking into consideration various tax incentives). 
 \begin{enumerate}
\item Make a table showing how much they will have after fifteen years if every month they contribute  \$100 vs.\ \$500 vs.\ \$\text{1,000}. Use the \textsc{Future Value Annuity Formula}. \vfill  \vfill
\item Tim's parents decide to put \$\text{15,000} into the account right now.  How much will that add to the value in fifteen years?  Use the \textsc{Compound Interest Formula}.  \vfill  \vfill
\end{enumerate}
%\item 
%\begin{tabular} {|c |c |c |c|}\hline
%$P$ & 100 & 500 & 1000 \\ \hline
%$A$ & 27,640.60 & 138,203.01 & 276,406.03 \\ \hline
%\end{tabular}

\item Use the \textsc{Equivalent APR Formula} to find the APR for each of the following published interest rates (compounded monthly) offered by recent credit card companies.
\bigskip

%\item (a) 9.38\% \quad (b) 12.58\% \quad (c) 22.17\%
\begin{enumerate}
\item 9\% \vfill
\item 12.8\%   \vfill
\item 20.19\% \vfill
\end{enumerate}

\newpage %%%%%%

\item  Cesar and Eliana are looking at three different houses to buy.  The first, a large new townhouse, for \$\text{240,000}.  The second, a small but charming bungalow, for \$\text{260,000}.  The third, a large 2-story house down the block, for \$\text{280,000}. 
%\item (a) 
%\begin{tabular} {|c |c |c |c|}\hline
%$A$ & 240,000 & 260,000 & 280,000  \\ \hline
%$P$ & 1,077.71 & 1,167.52 & 1,257.33 \\ \hline
%\end{tabular}
%\quad (b) Increases monthly payment by \$89.81
 \begin{enumerate}
\item Calculate the monthly payment for each house for a 30-year mortgage at 3.5\% interest compounded monthly.   Use the \textsc{Loan Payment Formula}. 
\bigskip

\begin{description}
\item[Townhouse] ~\vfill
\item [Bungalow] ~\vfill
\item [2-Story] ~\vfill
\end{description}
\item Describe the effect on Cesar and Eliana's monthly payment of each \$\text{20,000} increase in the house price  at this interest rate. \vfill
\end{enumerate}

\end{enumerate}

