\section{Solving power equations (and roots)}

%SU consider adding more homework problems that include formulas from geometry?

% SU do you want to have some solving inversely proportional to a power here?

There's an old saying -- ``when life gives you lemons, make lemonade.''  But how many lemons do you need? It turns out a reasonable equation describing the yield of lemonade from a single lemon is given by $$J = 0.0185C^3$$ where $J$ is the juice, measured in tablespoons, and $C$ is the circumference of the lemon, measured in inches.  (In case you've forgotten, the circumference is the distance \emph{around} the lemon. Think of taking a piece of string and wrapping it around the middle part of the lemon. Then lay the string on a rule to see how long it is.)

A small lemon might measure 6 inches in circumference.  According to our equation, it would yield $$J = 0.0185 (6)^3 = 0.0185 \times \underline{6} \wedge 3 = 3.996 \approx 4\text{ tablespoons}$$ A regular-sized lemon, say 8 inches in circumference, would yield $$J = 0.0185 (8)^3 = 0.0185 \times \underline{8} \wedge 3 = 9.472 \approx 9.5\text{ tablespoons}$$ %And a giant lemon, say 10 inches in circumference, would yield $$J = 0.0185 (10)^3 = 0.0185 \times \underline{10} \wedge 3 = 18.5\text{ tablespoons}$$

%By the way,16 tablespoons equals one cup of liquid so our giant lemon yields over a cup of juice.  Well, at least according to this equation.  

Let's make a table of values and look at a graph of this function.  We've added some values (including some unrealistic ones) to see the shape better.
\begin{center}
\begin{tabular} {|c |c |c |c |c |c |c |c |c|}\hline
$C$ & 0 & 2 & 4 & 6 & 7 & 8 & 9 & 10 \\ \hline
$J$ & 0 & 0.148 & 1.184 & 3.996 & 6.3455 & 9.472 & 13.4865 & 18.5 \\ \hline
\end{tabular}
\end{center}

SU GRAPH

\vspace{2in}

How large a lemon would yield half a cup of juice?  Remember 1 cup is 16 tablespoons so $$\frac{1}{2} \text{ cup} \ast \frac{16 \text{ tablespoons}}{1 \text{ cup}} = 16 \div 2 = 8 \text{ tablespoons}$$ From our graph it look like 7.5 inches in circumference should be pretty close.  Let's use successive approximation to get the answer to two decimal places. We'll round our value of $J$ to two decimal places to fit easily into the table.

\begin{center}
\begin{tabular} {|c |c |c |c |c |c |c |c |c|}\hline
$C$ & 7 & 8 &7.5 & 7.6 & 7.55 & 7.56 & 7.57 \\ \hline
$J$ & 6.35 & 9.47 & 7.80 & 8.12 & 7.96 & 7.99 & 8.03 \\ \hline
vs.\ 8 & low & high & low & high & low &not quite & just over \\ \hline
\end{tabular}
\end{center}

\noindent So a lemon with circumference of approximately 7.57 inches should yield about half a cup of juice.

That didn't take too much work.  But this chapter is all about solving equations.   Much as we have learned to love successive approximation (and for good reason as it works on any type of equation), you might be happy to know that there is a way to solve power equations exactly.  Here's how.  %SU this moves back to solving expl equations if you switch order.

Start with what we're looking for $$J = 8$$ Next, use our equation $J=0.0185C^3$ to get $$0.0185C^3=8$$ We want to find the value of $C$, so we can divide both sides by 0.0185 to get $$\frac{\cancel{0.0185}C^3}{\cancel{0.0185}}=\frac{8}{0.0185}=432.432\ldots$$  Thus we have $$C^3 = 432.432\ldots$$ 

How can we find $C$ just knowing $C^3$?  We take cube roots of both sides to get $$\sqrt[3]{C^3} = \sqrt[3]{432.432\ldots}= 432.432\ldots^{1/3}=432.432\ldots \wedge ( 1 \div 3)= 7.56204\ldots $$   As before we see a lemon of about 7.57 inches should yield half a cup of juice.

Perhaps a brief digression on roots is in order.  (And, that mysterious $\wedge (1\div3)$ business needs some explanation, I expect.) For example, we know $$2^3=2 \wedge 3=8$$ which tells us $$\sqrt[3]{8}=2$$ The \emph{\textbf{cube root}} of a number is whatever you would cube to get that number.  Similarly $$\sqrt[3]{1000} = 10$$ because $$10^3=10 \wedge 3 = 1000$$ 

Some calculators have a key, or sequence of keys, that will take the cube root.  For example, if your calculator has a key labeled $\sqrt[x]{y}$, then you can use it to find the cuberoot by doing $$\sqrt[3]{8}=3\text{ }\sqrt[x]{y}\text{ }8=2$$  Be careful if the key is just labelled $\sqrt{x}$ because that only takes square roots and can't be used for cube roots.

Otherwise you need to know that cube root is equivalent to raising to the 1/3rd power.  So you can calculate $$\sqrt[3]{8}=8^{1/3}=8\wedge(1\div3)=2$$ as we did in our example.  Notice how the 1/3 is in parentheses.  That's because we want the division before the exponent but the usual order of operations is the other way.  If your calculator has a fraction entering key, you might be able to do $$\sqrt[3]{8}=8^{1/3}=8\wedge1/3=2$$ instead, but don't forget to use the fraction key $/$ not the divides key $\div$.

In general, the \emph{\textbf{$n$th root}} of a number is whatever number you would raise to the $n$ power to get the number.  The formula is
\begin{center}
\textsc{The Root Formula} \\ ~\\
The equation $C^n=v$ has solution $C= \sqrt[n]{v} = v^{1/n}$
\end{center}

SU -- decide which comes first: solving expl or solving power, then have 2nd one refer to the 1st.

Back to our lemonade example.  Let's practice solving this type of equation.  If we wanted a lemon that yields 10 tablespoons of juice we would solve $$J=10$$ by putting in our equation we get $$0.0185C^3=10$$  Next, divide each side by 0.0185 to get $$\frac{\cancel{0.0185}C^3}{\cancel{0.0185}}=\frac{10}{0.0185}$$ so that $$C^3 = 10 \div 0.0185 = 540.540\ldots$$ Then take cuberoots to get $$C=\sqrt[3]{ 540.540\ldots} = 540.540\ldots\wedge(1 \div 3) = 8.14596\ldots \approx 8.1 \text{ inches}$$

As before we see that we solve in the reverse order of operations.
\begin{quote} If evaluating, first do $\wedge$ (power), then do $\times, \div$

If solving, first do $\times,\div$, then do $\sqrt[n]{\text{ }}$ (root)\end{quote}

A few pages of calculations into our example and we have only half a cup of lemonade to show for it. How about if we're trying to make a 1/2 gallon pitcher of lemonade?  Suppose the store sells lemons by the bag, where all the lemons in the bag are just about the same size (all small, all medium, etc.).  How many lemons of a fixed size will it take to make 1/2 gallon of lemonade?  We can write a new equation to describe this situation.

First, convert 1/2 gallon into tablespoons. $$\frac{1}{2} \text{ gallon } \ast \frac{4 \text{ quarts}}{1 \text{ gallon}} \ast \frac{4 \text{ cups}}{1 \text{ quart}} \ast \frac{16 \text{ tablespoons}}{1 \text{ cup}} = 4 \times 4 \times 16 \div 2 = 128 \text{ tablespoons}$$

Let $L$ be the number of lemons (of a fixed size) and suppose each lemon yields $J$ tablespoons of juice. For example, if $J$ = 10 tablespoons, then $$\frac{128 \text{ tablespoons}}{10 \frac{\text{tablespoons}}{\text{lemon}}} = 128 \div 10 = 12.8 \text{ lemons}$$ so we would need 13 lemons.  In general the number of lemons we need is given by the equation $$L=\frac{128}{J}$$ 

Look at a table of a few values and graph for this function.  Notice that it's decreasing because the larger the lemons, the fewer we need to use.  That makes sense, doesn't it?

\begin{center}
\begin{tabular} {|c |c |c |c |c |c|}\hline
$J$ & 4 & 6 & 10 & 12 &16 \\ \hline
$L$ & 32.0 & 21.3 & 12.8 & 10.7 & 8 \\ \hline
\end{tabular}
\end{center}

SU NEED GRAPH \vspace{2in}

Last time I made a pitcher of lemonade it took 9 lemons. How much juice did each yield (assuming they're all about the same size)?  Let's solve our equation to find the answer.  Start with what we want $$L=9$$ and use our equation to get $$\frac{128}{L}=9$$  We haven't seen an equation like this before, where the independent variable in in the denominator (bottom) of the fraction, but not to worry. Remembering that $\frac{128}{L}$ means $128 \div L$, we can multiply both sides of the equation by $L$ to get $$\cancel{L}\ast\frac{128}{\cancel{L}} = 9\ast L$$ so that $$9L=128$$  (We switched the variable onto the left-hand side for convenience.)  Now the equation looks much more familiar.  Dividing both sides by 9 gives us $$\frac{\cancel{9}L}{\cancel{9}}= \frac{128}{9}$$ so $$L=\frac{128}{9} = 128\div 9 = 14.22\ldots \approx 14.2 $$  Each lemon must have yielded around 14.2 tablespoons of juice.

And what goes better with lemonade than lemon cake.  For that we're going to need some grated lemon peel.  As with juice, the amount of lemon peel depends on the size of the lemon.  One equation is $$P=0.061C^2$$ where $C$ is the circumference of the lemon, measured in inches, as before and $P$ is the amount of lemon peel, measured in tablespoons.  For example, a lemon of circumference 7 inches will produce about 3 tablespoons of grated lemon peel because $$0.061\ast7^2 = 0.061 \times \underline{7} \wedge 2 = 2.989 \approx 3$$

As before, we can look at a table of select values (not all of which are realistic).

\begin{center}
\begin{tabular} {|c |c |c |c |c |c |c |c |c|}\hline
$C$ & 0 & 2 & 4 & 6 & 7 & 8 & 9 & 10 \\ \hline
$P$ & 0 & 0.244 & 0.976 & 2.196 & 2.989 & 3.904 & 4.941 & 6.1 \\ \hline
\end{tabular}
\end{center}

SU NEED GRAPH \vspace{2in}

What size lemon would give 4 tablespoons of grated lemon peel?  From the graph it looks like just over 8 inches in circumference should do.  A quick successive approximation shows it's around 8.1 inches in circumference.

\begin{center}
\begin{tabular} {|c |c |c |c |c |c |c |c |c|}\hline
$C$ & 8 & 9 &  8.1\\ \hline
$P$ & 3.904 & 4.941& 4.00221 \\ \hline
vs.\ 4 & bit low & high & close \\ \hline
\end{tabular}
\end{center}

To solve exactly we begin with what we want $$P=4$$ and use our equation to get $$0.061C^2=4$$ Then divide both sides by 0.061 to get $$\frac{\cancel{0.061}C^2}{\cancel{0.061}}= \frac{4}{0.061}$$ which simplifies to $$C^2 = \frac{4}{0.061} = 4 \div 0.061 = 65.57\ldots $$ Now to undo the square we use regular square roots.  There is most likely a special key on your calculator for square roots, but we'll do it the same as any other root here. $$\sqrt{C^2} = \sqrt{65.57\ldots} $$ so $$C = \sqrt{65.57\ldots} = 65.57\ldots^{1/2} = 65.57\ldots \wedge ( 1 \div 2) = 8.097\dots \approx 8.1$$ as before. 

%\newpage

%\input{Solving_power_equationsEX.tex}