%\section{Solving linear inequalities}

 \input{GoDoWS}

\noindent \textbf{Do you know \ldots}

\begin{itemize}
\item Common phrases that indicate an inequality? 
\item How to represent the idea of ``between'' using a double-sided inequality? 
\item Why we ``do the same thing to each side'' of an inequality when solving? 
\item How to solve a linear inequality? A chain of inequalities?
\item Why the inequality sign is reversed if we switch sides of the equation? 
\item When to solve an inequality, as opposed to solving an equation? 
 \input{NotSure}
\end{itemize}

\subsection*{Exercises}

\begin{enumerate} 
\setcounter{enumi}{4}

\item  Recall that the conversion between Fahrenheit ($F$) and Celsius ($C$) temperatures is given by the equation $$F = 1.8C+32$$
\begin{enumerate}
\item Evaluate the equation at the appropriate values to check that $0^\circ$C$=32^\circ$F, $20^\circ$C$=68^\circ$F, and $37^\circ$C$=98.6^\circ$F.
\item Set up and solve an equation to find the Celsius equivalent of brownie-baking temperature ($350^\circ$F).
\item You're planning a trip to Norway over Christmas and have heard it's will be around 10$^\circ$C.  What sort of jacket will you need?  Convert to Fahrenheit to decide. 
\item You want to explain to your Norwegian hosts that back in Minnesota this time of year temperatures can range between -20$^\circ$F and 40$^\circ$F.  Express this range in Celsius instead.  Set up and solve a chain of inequalities.
\item Your Norwegian hosts ask about the temperature in Minnesota during the summer.  You explain that summer temperatures typically range from 55$^\circ$F and 105$^\circ$F.  Express this range in Celsius instead.   Set up and solve a chain of inequalities.
\end{enumerate}

\item After borrowing some money through a line of credit on my bank account, I started paying off the interest plus \$250 a month.  Even once the loan is paid off I plan to continue to deposit \$250 each month to start savins.  That means my balance is given by the equation
$$A = 250M- \text{2,189.57}$$
where $M$ is the number of months since the loan and $A$ is the account balance, in dollars. \hfill \emph{Story also appears in 2.1 \#3}
\begin{enumerate}
\item Set up and solve an inequality to determine when I will have paid off my line of credit. That means the account balance will be \$0 or more.
\item Set up and solve an inequality to determine when I will have saved at least \$2,000.
\end{enumerate}  

\item When Gretchen walks on her treadmill, the number of calories per mile that she burns is given by the equation $$C=125M$$ where $C$ is the number of calories Gretchen burns by walking $M$ miles.  

\hfill \emph{Story also appears in 2.1 and 3.1 Exercises}
\begin{enumerate}
\item Draw a graph showing the number of calories Gretchen burns if she walks 0, 1, 2, or 3 miles.
\item How far does she need to walk to burn at least 200 calories? Set up and solve an inequality.
\item Highlight the part of your graph where she burns at least 200 calories.
\end{enumerate}

\item The water in the local reservoir has been dropping steadily.  In fact, $$D = 47-1.5W$$ where $D$ is the depth of the water (in feet) after $W$ weeks.  Any depth below 20 feet is considered dangerously low.  When will that happen, assuming no change in the weather?  Set up and solve an inequality.  And, check your answer.

\hfill \emph{Story also appears in 2.1 \#2 and 4.1 \#3}

\item A manufacturer makes family-sized bags of potato chips, advertised as containing 200 grams each.  In fact it's difficult to control the exact weight of a bag of potato chips, so it varies. The standard deviation is rather high, about 3.8 grams per bag.  The company would rather have bags too heavy than too light, lest they be accused of false advertising, so their average bag actually weighs 207 grams.  It turns out that approximately 97\% of all bags of chips weigh 200 grams or more.  We can compute the standard $Z$-score of a given bag of chips weighing $B$ grams using the equation $$Z = \frac{B-207}{3.8}$$
\begin{enumerate}
\item What is the $Z$-score for a bag of potato chips weighing the advertised 200 grams?  Remember above average $Z$-scores are positive and below average $Z$-scores are negative, so your answer should be negative.  
\item About \nicefrac{3}{4} of all bags of chips will have $Z \ge -.67$.  What weight bag has $Z$-score of $-.67$?  Set up and solve an inequality.
\item A standard $Z$-score between -1 and +1 is considered ordinary.  What weight bags are considered ordinary? 
\item Oh, and if a serving size is 28 grams (approximately 1 ounce), how many servings are in a bag that weights 207 grams?
\end{enumerate}

\item  The cost of vacation to Cork, Ireland from the Minneapolis/St. Paul airport for two people is given by the equation formula $$C = \text{2,828} + 310N$$ where $C$ is the total cost in U.S. dollars and $N$ is the number of days.  Ciara wants to take her boyfriend Seamus to Cork to meet Ciara's grandmother.
\begin{enumerate}
\item What would it cost Ciara to go to Ireland with Seamus for six days?
\item What might the number \text{2,828} mean in terms of the story, and what are its units?
\item What might the number 310 mean in terms of the story, and what are its units?
\item  If Ciara has budgeted up to \$\text{10,000}, how many days can they afford to spend in Ireland? Set up and solve an inequality to find the answer.
\end{enumerate}

\end{enumerate}


