\chapter{Answers to exercises}

These answers to the exercises are provided for you to check your work.

\begin{itemize}
\item Ask your instructor about access to the solutions of the Practice Exercises \#1-4.
\item Graphs are not provided, so be sure to compare your graphs with those of a classmate.
\item Remember to use whatever method is indicated in the Exercises.
\end{itemize}

\noindent \textbf{When you're done \ldots}

\begin{itemize}
\item Not sure if your answers are close enough? Compare with a classmate or ask the instructor.  
\item Getting the wrong answers or stuck on a problem?  Re-read the section and try the problem again.   If you're still stuck, work with a classmate or go to your instructor's office hours.
\item It's normal to find some parts of some problems difficult, but if all the problems are giving you grief, be sure to talk with your instructor or advisor about it.  They might be able to suggest strategies or support services that can help you succeed.
\item Make a list of key ideas or processes to remember from the section.  The ``Do you know?'' questions can be a good starting point.
\end{itemize}

\newpage  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Variables} % Chapter 1

\bigskip

\subsection{Variables and functions -- Answers to exercises} %1.1

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % It's about time
\begin{enumerate}
\item about 17 years ago % Nussbaum's walnut tree
\item .058, .084 % Stephen's BAC
\item 20 mph % McKenna highway exists
\item in 45 days, early November \\ (ignores daylight savings) % sunset time
\end{enumerate}

\item % 40 deg temp drop 3/hr
\begin{enumerate}
\item 3
\item $T = $ temperature ($^\circ$F) $\sim$ dep $H = $ time (hours since noon) $\sim$ indep
\item 2:40 p.m.
\end{enumerate}

\item % Mrs. Nystrom's social security benefit
\begin{enumerate}
\item $S =$ Mrs.\ Nystrom's social security benefit (\$) $\sim$ dep \\ $Y=$ time (years since 2009) $\sim$ indep
\item \$839.34
\item $\approx$ 2014
\end{enumerate}

%\columnbreak

\item % postage stamps
\begin{enumerate}
\item $S=$ number of stamps $\sim$ indep \\ $C=$ total cost (\$) $\sim$ dep
\item $C$ is dependent, $S$ is independent
\item 21 stamps 
\end{enumerate}

\item % Sofia's car
\begin{enumerate}
\item $M=$ car's mileage (thousands of miles) $\sim$ indep \\ $C=$ value of Sof\'ia's car (\$) $\sim$ dep
\item $C$ is a function of $M$
\item maybe $0 \le M \le 250,000$
\item $\approx$ another 5 years
\end{enumerate}

\item 
\begin{enumerate}
\item $L=$ intensity of light (\%) $\sim$ dep \\ $F$ = distance (feet) $\sim$ indep % lamplight
\item $M=$ time (minutes) $\sim$ dep \\$F=$ thickness (inches) $\sim$ indep  % Mrs. Weber fish
\item $W =$ electricity (watts) $\sim$ dep \\ $S=$ wind speed (mph) $\sim$ indep % wind turbine
\end{enumerate}
\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Tables and graphs -- Answers to exercises} %1.2

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item %Earth's pop
\begin{enumerate}
\item 1.26 billion
\item $\approx$1.35 billion
\item $\approx$ 3 billion
\item $\approx$ 1930
\item $\approx$ 2015
\item $Y=$ time (years since 1800) $\sim$ indep \\ $P=$ population of Earth (billions of people) $\sim$ dep
\end{enumerate}

\item % truck rental
\begin{enumerate}
\item \$55.00
\item 200 miles
\item $\approx$ \$46
\item $\approx$ \$20
\item $\approx$ 120 miles
\item $D=$ distance (miles) $\sim$ indep \\ $R=$ rental cost (\$) $\sim$ dep \\ maybe $0 \le D \le 300$
\item Graph
\item $\checkmark$
\end{enumerate}

\item % 40 deg temp drop 3/hr
\begin{enumerate}
\item 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$H$ & 0 & 1 & 2& 3 & 4 & 10 \\ \hline
$F$ & 40 & 37 & 34 & 31 & 28 & 10  \\ \hline
\end{tabular}
\item Graph
\item 2:40 p.m.
\item $\approx$ midnight, unusual but realistic for the midwest
\end{enumerate}

\item % Mrs. Nystrom's social security benefit
\begin{enumerate}
\item  See \textbf{Table A}, below.
\item Graph
\item in 2014
\item $\approx$ \$1,100
\end{enumerate}

\item % heat index
\begin{enumerate}
\item 100$^\circ$F
\item $\approx$ 66\%
\item $\approx$ 78\%
\item maybe $78\% \le H \le 100\%$
\item $\approx$ 120$^\circ$F
\item $H=$ humidity (\%) $\sim$ indep \\ $F=$ heat index ($^\circ$F) $\sim$ dep \\
 $0 \le H \le 100$
 \item Graph
 \item $\checkmark$
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip
\begin{description}
\item[\quad Table A: \quad] \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
year & 2009 & 2010 & 2011 & 2012 & 2013& 2014 & 2015 \\ \hline
$Y$ & 0 & 1 & 2& 3 & 4 & 5 & 6 \\ \hline
$S$ & 746.17 & 776.02 & 807.06 & 839.34 & 872.91 & 907.83 & 944.14 \\ \hline
\end{tabular}
\end{description}
\bigskip

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Rate of change -- Answers to exercises} % 1.3

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % diver, revisited
\begin{enumerate}
\item $\checkmark$
\item $\approx$ 8.8 feet per second
\end{enumerate}

\item % truck rental
\begin{enumerate}
\item \$.35/mile,  \$.35/mile,  \$.35/mile
\item \$44.50
\item \$20
\end{enumerate}

\item %wind turbine
\begin{enumerate}
\item $W =$ electricity (watts) $\sim$ dep \\ $S=$ wind speed (mph) $\sim$ indep 
\item Graph
\item 240 watts/mph
\item 1,680 watts/mph, increasing faster
\item  4,560 watts/mph, fastest yet
\end{enumerate}

\item % Earth's pop
1990-2000

\item % backpacks
\begin{enumerate}
\item $C=$ cost per backpack (\$/backpack) $\sim$ dep \\ $B$ = number of backpacks $\sim$ indep \\
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$B$ & 10 & 40 & 70 & 100\\ \hline
$C$ & 39 & 18 & 15 & 30  \\ \hline
\end{tabular}
\item -\$.70/backpack, -\$.10/backpack, +\$.50/backpack
\item maybe $0 \le B \le 60$
\item need more space, another machine, overtime pay?
\item Graph
\item $\approx 60$ 

\end{enumerate}

\item %Paloma's beach
\begin{enumerate}
\item -3.75 feet/year, -3.75 feet/year, -3.75 feet/year
\item Erosion makes the beach less deep.
\item 3.75 feet/year
\item Graph
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Units -- Answers to exercises} % 1.4

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Michael Phelps
\begin{enumerate}
\item $\approx 1.913$ minutes
\item $\approx 104.5$ meters/min
\item $\approx 3.9$ mph
\item 1.9 minutes
\item $\approx 105.3$ meters/min
\item $\approx 3.9$ mph
\end{enumerate}

\item \begin{enumerate}
\item 44 pounds % suitcase Africa
\item $\approx 17$ servings % cookies
\item just over 5 cans % thirst quencher pop
\end{enumerate}

\item \begin{enumerate}
\item 1,320 yards % football
\item 1.125 feet % fabric
\item 4.2 inches % basil plant
\end{enumerate}

\item % pollution spill
\begin{enumerate}
\item See \textbf{Table B}, below.
\item 3 hours and 35 minutes
\end{enumerate}

\item %heart beats a year
$\approx$ 700,000 heartbeats/week \\ $\approx 37$ million heartbeats/year \\
(calculated  705,600 and 36,792,000)

\item %Harold's textbook says
\begin{enumerate}
\item $\approx 24.5$ mph
\item $\approx 9$ square miles
\item $\approx \$2.38$/gallon
\end{enumerate}

\end{enumerate}
\end{multicols}

 \bigskip
\begin{description}
\item[\quad Table B: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
Day & Mon & Tue & Wed & Thu & Fri \\ \hline
Minutes & 35 & 71 & 15 & 0 & 98\\ \hline
Hours & .583 & 1.183 & .25 &0 & 1.683\\ \hline
H:MM & 0:35 & 1:11 & 0:15 & 0:00 & 1:38 \\ \hline
\end{tabular}
\end{description}
\bigskip 

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Metric prefixes and scientific notation -- Answers to exercises} % 1.5

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % computer memory, revisisted
\begin{enumerate}
\item $\approx 24$ files
\item just over 1 hour
\item $\approx$ 5.5 KB
\end{enumerate}

\item % Billy Bob
\begin{enumerate}
\item $\approx 11$ days, $\approx 32$ years, $\approx 32,000$ years 
\item 32 years old
\item \emph{Hint: use 31 years, 259 days, 1 hour, and 47 minutes from the day and time of your birth.  Leap years have an extra day:  2012, 2008, 2004, 2000, 1996, \ldots}
\end{enumerate}

\item % Proton electron
\begin{enumerate}
\item See below
\item proton
\item about 1,800 times heavier
\item $\approx 1.69 \times 10^{25}$
\end{enumerate}

\item %servings
\begin{enumerate}
\item 8 servings
\item $\approx 5$ servings
\end{enumerate}

\item %Rayka's cells
\begin{enumerate}
\item $\approx 6.4 \times 10^{19}$ cells
\item $\approx 64$ quintillion cells
\end{enumerate}

\item % body mass index
\begin{enumerate}
\item $\checkmark$
\item Yes
\item $\approx 25.2$ BMI, no
\item ``Jared, you're right on the upper limit for normal BMI.''
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{description}
\item[\quad proton \quad] .000~000~000~000~000~000~000~000~001~672~62 
\item[\quad electron \quad] .000~000~000~000~000~000~000~000~000~000~910~938
\end{description}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

%\newpage  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Equations} % Chapter 2

\bigskip

\subsection{A first look at linear equations -- Answers to exercises} % 2.1

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % plumbers, revisited
\begin{enumerate}
\item $P=50+120T$
\item $P=45+55T$
\item $P=90T$
\item $P=150$
\end{enumerate}

\item %Abdulwali's restaurant
\begin{enumerate}
\item No
\item $A=$ Abduwali's profit (\$) $\sim$ dep \\ $M=$ time (months) $\sim$ indep \\ $A = 6,300M-82,500$
\item Slope is \$6,300/month; it's how much money he earns in a month.\\  Intercept is -\$82,500; it's his start up costs.
\item See \textbf{Table C}, below.
\end{enumerate}

\item %Gretchen treadmill
\begin{enumerate}
\item 287.5 calories
\item $C=$ energy (calories) $\sim$ dep \\ $M=$ distance (miles) $\sim$ indep \\ $C=125M$
\item Slope is 125 cal/mile; it's the rate at which Gretchen burns calories. \\ Intercept is 0; it's how many calories she burns walking 0 miles.
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$M$ & 0 & 1 & 2 & 3 & 4 \\ \hline
$C$& 125 & 250 & 375 & 500 & 625  \\ \hline
\end{tabular}
\end{enumerate}

\item % burger discount
\begin{enumerate}
\item $H=$ daily high temperature ($^\circ$F) $\sim$ indep \\ $B =$ price bacon double cheeseburger (\$) $\sim$ dep \\ $B=7.16-.02H$
\item decreasing
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$H$ & 65 & 75 & 90 \\ \hline
$B$ & 5.86 & 5.66 & 5.36  \\ \hline
\end{tabular}
\item Graph
\end{enumerate}

\item % hospital beds
\begin{enumerate}
\item $B=$ number of beds $\sim$ dep \\ $Y=$ time (years since 1975) $\sim$ indep
\item  $B = 1,466,000-16,000Y$
\item decreasing
\item See \textbf{Table D}, below.
\item guess in 2040? actually in 2036
\end{enumerate}

\item % traffic
\begin{enumerate}
\item $C=$ traffic (cars/hour)$\sim$ dep \\ $W=$ time (weeks) $\sim$ indep \\ $C = 1,450 + 130W$
\item See \textbf{Table E}, below.
\item just over 4 years
\item slightly over 8 years
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip
\begin{description}
\item[\quad Table C: \quad]  \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$M$ & 0 & 1 & 6 & 10 & 12 & 24\\ \hline
$A$ & -82,500 & -76,200 &  -44,700 & -19,500 & -6,900 & 68,700\\ \hline
\end{tabular}
\end{description}
 
\begin{description}
\item[\quad Table D: \quad]  \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
year & 1980 & 1990 & 2000 & 2010 & 2020\\ \hline
$Y$ & 5 & 15 & 25 & 35 & 45  \\ \hline
$B$ & 1,386,000 & 1,226,000 & 1,066,000 & 906,000 & 746,000 \\ \hline
\end{tabular}
\end{description}

\begin{description}
\item[\quad Table E: \quad] \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$W$ & 0 & 2 & 4 & 6 & 8 & 10\\ \hline
$C$ & 1,450 & 1,710 & 1,970 & 2,230 & 2,490 & 2,750  \\ \hline
\end{tabular}
\end{description}
\bigskip

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection{A first look at exponential equations -- Answers to exercises} % 2.2

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Mai's salary raise
\begin{enumerate}
\item \$83,920.20
\item 7.59\%
\item same
\item 6\% first because get higher salary in year 2
\end{enumerate}

\item % school children single parent HH
\begin{enumerate}
\item 1.03
\item 1,495 children
\item $Y=$ time (years since 2010) $\sim$ indep \\ $C=$ number of children from single parent households $\sim$ dep \\ $C = 1,290 \ast 1.03^Y$
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
year & 2010 & 2015 & 2020 & 2030\\ \hline
$Y$ & 0 & 5 & 10 & 20 \\ \hline
$C$  & 1,290 & 1,495 & 1,733 & 2,329 \\ \hline
\end{tabular}
\item Graph
\end{enumerate}

\item % Um Archivo earnings
\begin{enumerate}
\item $U=$ earnings (\$ billions) $\sim$ dep \\ $Y=$ time (years since 2012) $\sim$ indep \\ $Y = 42.7 \ast 1.17^Y$
\item $\approx$ \$149.9 billion
\item 17\% is too high; they earned much less than expected.
\item Graph
\end{enumerate}

\item % poultry
\begin{enumerate}
\item $P=$ poultry population (millions of tons) $\sim$ dep \\ $Y =$ time (years since 2005) $\sim$ indep \\$P = 78 \ast 1.016^Y$
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
year & 2005 & 2010 & 2020 & 2050\\ \hline
$Y$ & 0 & 5 & 15 & 45 \\ \hline
$P$ & 78 & 84.4 & 99.0 & 159.3 \\ \hline
\end{tabular}
\end{enumerate}

\item % ebook sales
\begin{enumerate}
\item $E=$ ebook sales (\$ million) $\sim$ dep \\ $M=$ time (months since Jan. 2008) $\sim$ indep 
\item 1.063
\item $E = 5.1 \ast 1.063^Y$
\item It's just a little lower than the equation predicted ($\approx$ \$22.1 million)
\item \$414.9 million
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Using equations -- Answers to exercises} % 2.3

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Chevy Tahoe
\begin{enumerate}
\item 398.6 feet
\item 234.6 feet
\item No
\item No, 146.7 feet behind car but it takes 183.2 feet to stop.  Crash.
\end{enumerate}

\item % lamplight
\begin{enumerate}
\item 100\%, 25\%, $\approx$ 11\%
\item $\approx$ 20\%
\item $\approx -21\%$  so the light intensity drops about 21\%/foot  in the 2 foot range
\item Graph
\end{enumerate}

\item % urban garden
\begin{enumerate}
\item .25 feet
\item 344 square feet
\item 86 cubic feet
\item $\checkmark$
\item 39 cubic feet, 61.5 cubic feet, 73.5 cubic feet
\end{enumerate}

\item % newspapers
\begin{enumerate}
\item 1,945 newspapers \\ 1,805 newspapers \\ 1,594 newspapers \\ 1,344 newspapers.
\item 1995-2010
\item Graph
\end{enumerate}

\item % Mrs. Weber fish
\begin{enumerate}
\item $\approx$ 8 minutes, $\approx$ 12 minutes,$\approx$ 16 minutes,$\approx$ 20 minutes
\item It takes about 8 minutes to grill a piece of fish that's only .25 inch.  For a piece of fish 1 inch thick it takes about 12 minutes.  For 1.5 inches it takes about 16 minutes and for 2 inches it takes about 20 minutes.
\item Graph
\item The thicker the piece of fish, the longer it takes to grill.
\end{enumerate}

\item % Wynter's wages
\begin{enumerate}
\item \$15.60/hour, \$13.00/hour
\item $E=$ hourly wage earned (\$/hour) $\sim$ dep \\ $H=$ time (hours) $\sim$ indep \\$\displaystyle E = \frac{780}{H}$
\item The more hours he works in a week, the less he is (effectively) paid per hour.
\end{enumerate} 

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Approximating solutions of equations -- Answers to exercises} % 2.4

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % population China
\begin{enumerate}
\item  in 2035
\item  in 2015
\item start $= 1.343$ billion and  \\ growth factor $= 1.0131$
\item Graph
\end{enumerate}

\item % going out of business
\begin{enumerate}
\item \$104.4 billion,  \$128 billion, \$94.5 billion
\item in 2010
\item  in 2008
\end{enumerate}

\item % thick glass
\begin{enumerate}
\item Each thickness is one inch.  \\ 75\% of 56.25\% $= .75 \times 56.25 = 42.1875 \approx 42.19\%$
\item 3 inches thick, $\approx 31.64\%$
\item $L=$ light (\%) $\sim$ dep \\ $W=$ thickness of window glass (inches) $\sim$ indep \\ $L = 100 \ast .75^W$
\item $\approx$ 8 inches
\item Graph
\end{enumerate}

\item % wind turbines
\begin{enumerate}
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$S$ & 10 & 25 & 40 \\ \hline
$W$ & 2,400 & 37,500 & 153,600  \\ \hline
\end{tabular}
\item Graph
\item $\approx$ 17.4 mph
\end{enumerate}

\item % Stephen's BAC
\begin{enumerate}
\item See \textbf{Table F}, below. 
\item Graph
\item in about 2\nicefrac{1}{2} hours
\end{enumerate}

\item % urban garden
\begin{enumerate}
\item 2.94 feet
\item $\approx$ 2'11" Yes
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip
\begin{description}
\item[\quad Table F: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$H$ & 0 & 1 & 2 & 3 & 4\\ \hline
$S$ & .040 & .058 & .084 & .122 & .177 \\ \hline
\end{tabular}
\end{description}
\bigskip

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Finance formulas -- Answers to exercises} % 2.5

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Hector's finances, revisited
\begin{enumerate}
\item $\checkmark$
\item $\checkmark$
\item $\checkmark$
\item $\approx$ 4.594 APR
\end{enumerate}

\item % Ayah's savings
\begin{enumerate}
\item  \$41,883.82, \$48,610.75, \$37,051.28
\item $R=$ interest rate $\sim$ indep \\ $A=$ Ayah's account balance (\$) $\sim$ dep \\
See \textbf{Table G}, below.  Graph
\end{enumerate}

\item % Lue family mortgage
\begin{enumerate}
\item \$1,024.74/month
\item \$875.64/month
\item \$1,370.20/month
\end{enumerate}

\item % Kurt's CD
\begin{enumerate}
\item See \textbf{Table H}, below.  
\item $Y=$ time (years) $\sim$ indep \\ $K=$ Kurt's balance (\$) $\sim$ dep \\Graph
\end{enumerate}

\item % Soo Jin's college loans
\begin{enumerate}
\item 3.45\% APR
\item 8.19\% APR
\item 21.70\% APR
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip
\begin{description}
\item[\quad Table G: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$R$ & 0 & 1.9 & 6 & 11\\ \hline
$A$ & 35,000 & 41,883.82 & \$48,610.75 &\$37,051.28  \\ \hline
\end{tabular}
\end{description}
 
\begin{description}
\item[\quad Table H: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$Y$ & 0 & 1 & 5 & 12\\ \hline
$K$ & \$50,000 & \$52,437.84 & \$63,437.19 & \$88,525.30  \\ \hline
\end{tabular}
\end{description}
 \bigskip
 
\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Solving equations} % Chapter 3

\bigskip

\subsection {Solving linear equations -- Answers to exercises} % 3.1

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % charter boat
\begin{enumerate}
\item  \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$P$ & 0 & 4 & 10 & 20  \\ \hline
$C$ & 135 & 182.80 & 254.50 & 374\\ \hline
\end{tabular}
\item Cost to rent the boat, in \$
\item Cost per passenger, in \$/person
\item 16 passengers
\item Graph $\checkmark$
\end{enumerate}

\item  % Abdulwali's restaurant
\begin{enumerate}
\item 14 months
\item 29 months
\end{enumerate}

\item % postage stamps
\begin{enumerate}
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$S$ & 5 & 10 & 20\\ \hline
$C$ & 2.95 & 5.15 & 9.55 \\ \hline
\end{tabular}
\item $S=$ number of stamps $\sim$ indep \\ $C=$ total cost (\$) $\sim$ dep \\ $C = .44S + .75S$
\item 15 stamps
\item 21 stamps
\end{enumerate}

\item % Gretchen treadmill
\begin{enumerate}
\item 2.4 miles
\item $\approx$ 43 minutes
\item $\approx$ 7.3 miles
\end{enumerate}

\item % demand strawberries
\begin{enumerate}
\item 145 pints
\item See \textbf{Table I}, below.  
\item Graph
\item \$1.30/pint
\end{enumerate}

\item % traffic
\begin{enumerate}
\item in just over 4 weeks
\item in just over 8 weeks
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip
\begin{description}
\item[\quad Table I: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$P$ & 2.00 & 2.25 & 2.50 & 2.75 & 3.00 & 3.25 & 3.50\\ \hline
$D$ & 205 & 192 & 180 & 167 & 155 & 142 & 130  \\ \hline
\end{tabular}
\end{description}
\bigskip

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Solving linear inequalities -- Answers to exercises} % 3.2

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Fahrenheit and Celsius, revisited
\begin{enumerate}
\item $\checkmark$
\item 177$^\circ$F
\item 50$^\circ$F, sweatshirt
\item between $-29^\circ$C and $-4^\circ$C
\item between $13^\circ$C and $41^\circ$C
\end{enumerate}

\item % line of credit
\begin{enumerate}
\item in 9 months
\item in 17 months
\end{enumerate}

\item % Gretchen treadmill
\begin{enumerate}
\item Graph
\item 1.6 miles
\item See graph
\end{enumerate}

\item % reservoir
in 18 weeks

\item % potato chips
\begin{enumerate}
\item $\approx$ -1.84
\item $\approx$ 204.5 grams
\item between 203.2 and 210.8 grams
\item just over 7 servings
\end{enumerate}

\item % Cork travel
\begin{enumerate}
\item \$4,688
\item the cost of airfare for Ciara and Seamus, in \$
\item the cost of hotel or other daily expenses, in \$/day 
\item 23 days
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Solving power equations (and roots) -- Answers to exercises} % 3.3

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % lemon zest, revisited
\begin{enumerate}
\item Graph
\item $\approx 7\nicefrac{1}{2}$ inches
\item $\approx$ 15 inches
\end{enumerate}

\item % wind turbine
\begin{enumerate}
\item 17.4 mph
\item 26.6 mph
\end{enumerate}

\item % lamplight
\begin{enumerate}
\item $\approx$ 1.2 feet, or 1'2"
\item $\approx$ 1.7 feet, or 1'8"
\end{enumerate}

\item % Rodney's fish
\begin{enumerate}
\item $g \approx 1.223$ so 22.3\%
\item $g \approx 1.259$ so 25.9\%
\end{enumerate}

\item % rock from a high place
\begin{enumerate}
\item 64 feet, 256 feet
\item from $T=2$ to $T=4$
\item No
\item 400 feet, which really means it had hit the ground already
\end{enumerate}

\item % Wynter's wages
\begin{enumerate}
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$H$ & 40 & 50 & 60\\ \hline
$E$ &  19.50 & 15.60 & 13.00 \\ \hline
\end{tabular}
\item $\approx$ 84 hours/week
\item $\approx$ 56 hours/week
\item Graph
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Solving exponential equations (and logs) -- Answers to exercises} % 3.4

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % employee health insurance
\begin{enumerate}
\item $E=$ employee-paid cost of health insurance (\$/month) $\sim$ dep \\ $Y=$ time (years since 2003) $\sim$ indep \\ $E= 420 \ast 1.07^Y$
\item in 2007
\item in 2009
\item Graph
\end{enumerate}

\item % school children single parent HH
\begin{enumerate}
\item in 2039
\item in 2039 (again)
\item in 2044
\end{enumerate}

\item % thick glass
\begin{enumerate}
\item $\approx$ 8 inches
\item $\approx$ 2.4 inches
\item $\checkmark$
\end{enumerate}

\item % poultry
\begin{enumerate}
\item in 2017
\item in 2032
\end{enumerate}

\item % Darcy's hair color
\begin{enumerate}
\item about 8 washes
\item about 28 washes (or 1 month)
\item Graph
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Solving quadratic equations -- Answers to exercises} % 3.5

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Claude juggles, revisited
\begin{enumerate}
\item $\approx$ .17 seconds, .77 seconds\\
once going up, the other down
\item The calculator says ERROR because there's a square root of a negative number.  The max height of the beanbag was 6'6", so it is never 8' up.
\item 6.0625 feet, just over 6'
\end{enumerate}

\item % Cadillac Escalade, revisited
\begin{enumerate}
\item $\approx$ 52 mph
\item $\approx$ 52 mph (again)
\end{enumerate}

\item % backpacks
\begin{enumerate}
\item 35 backpacks
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$B$ & 10 & 40 & 70 & 100\\ \hline
$C$ & 39 & 18 & 15 & 30  \\ \hline
\end{tabular}

Graph, Yes
\item \$14/backpack
\end{enumerate}

\item % Mrs. Weber fish
\begin{enumerate}
\item about 2 inches
\item about .6 inches
\end{enumerate}

\item % going out of business
\begin{enumerate}
\item in 2010
\item in 2008
\item in 2000
\end{enumerate}

\item %kangaroo
\begin{enumerate}
\item 10.7272 feet, 9.2128 feet
\item $\approx 11.3852$ feet
\item just under 2 seconds (1.581 seconds)
\item in just over 1 seond (1.065 seconds)
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{A closer look at linear equations} % Chapter 4

\bigskip

\subsection {Modeling with linear equations -- Answers to exercises} % 4.1

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % affordable housing
\begin{enumerate}
\item $A=64,100+6,200Y$ \\ 14 years
\item $A=48,700+6,200Y$ \\ 17 years
\item 8,442 apartment/year
\end{enumerate}

\item % tuition per credit hour
\begin{enumerate}
\item $T=$ total tuition (\$) $\sim$ dep \\ $C=$ number of credit hours $\sim$ indep \\ $T = 870C+560$
\item Slope is \$870/credit; it's how much each credit hour costs.
\item Intercept is \$560; it's the one-time fee.
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$C$ & 3 & 12 & 16 \\ \hline
$T$ & 3,170 & 11,000 & 14,480  \\ \hline
\end{tabular}
\item $T=415C$
\item Slope is \$415/credit; it's how much each credit hour costs.
\item Intercept is \$0 because there's no fee.
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$C$ & 3 & 12 & 16 \\ \hline
$T$ & 1,245 & 4,980 & 6,640  \\ \hline
\end{tabular}
\item Graph
\item The line for the community college tuition is always below the line for the state university tuition.
\end{enumerate}

\item \begin{enumerate}
\item % Ahmed's virbunum
linear:  shrub grows at constant rate of .4 inches/week.
\item % rose gold
nonlinear:  increases but less and less as add more gold.
\item % Sea-ice
(probably) linear: sea-ice is decreasing around .043 millions of square miles per year.  
\item % wild rice
nonlinear: it's increasing, then decreasing, then increasing \ldots
\end{enumerate}

\item % 40 deg temp drop 3/hr
\begin{enumerate}
\item 2:40 p.m.
\item 1:20 a.m.
\end{enumerate}

\item % Shanille collects books
\begin{enumerate}
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$M$ & 0 & 1 & 12 & 36 \\ \hline
$B$ & 382 & 385 & 418 & 490 \\ \hline
\end{tabular}
\item $B=$ size of Shanille's collection (book) $\sim$ dep \\ $M=$ time (months) $\sim$ indep \\ $B = 382 + 3M$
\item just over 17 years
\item Graph, $\checkmark$
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Systems of linear equations -- Answers to exercises} % 4.2

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Quia Xun's locks
\begin{enumerate}
\item $E=1.55L$
\item \$10,667 locks
\item \$7,200 locks
\end{enumerate}

\item % washing floor
\begin{enumerate}
\item $T=$ time to wash floor (seconds) $\sim$ dep \\ $S=$ size of floor (square feet) $\sim$ indep \\ \textbf{rag:} $T = 80S$ \\ \textbf{mop:} $T=180+75S$
\item 36 squarefeet
\item bathroom: use rag \\ kitchen \& laundry room: use mop
\end{enumerate}

\item % Maria light bulbs
\begin{enumerate}
\item $M=$ time (months) $\sim$ indep \\ $C=$ total cost (\$) $\sim$ dep \\ \textbf{CFL bulb:} $C = 1+.95M$ \\ \textbf{LED bulb:} $24 + .60M$
\item See \textbf{Table J}, below.
\item Graph
\item 5\nicefrac{1}{2} years
\item 5\nicefrac{1}{2} years
\end{enumerate}

\item % exercise classes
\begin{enumerate}
\item See \textbf{Table K}, below.  Graph
\item $\approx 27$ exercise classes
\item 26 exercise classes
\item 26 exercise classes
\item If you're going to take at least 26 classes, it is worth it to buy the membership.  That's just 2 times a week for about 3 months.
\end{enumerate}

\item % supply and demand corn 
\begin{enumerate}
\item surplus at \$3.25/dozen \\ shortage at \$1.75/dozen
\item \$2.08/dozen
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$P$ & 1 & 2 & 3 & 4 \\ \hline
$S$ & 7.5 & 9 & 19.5 & 12 \\ \hline
$D$ & 10.1 & 9.2 & 8.3 & 7.4 \\ \hline
\end{tabular}

Graph, $\checkmark$
\end{enumerate}

\item % solar vs gas heat
\begin{enumerate}
\item 33 years 
\item 20 years
\item \$9,750
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip

\begin{description}
\item[\quad Table J: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$M$ & 1 & 6 & 18 & 36 \\ \hline
$C$ (CFL bulb) & 1.95 & 6.70 & 18.10 & 35.20\\ \hline
$C$ (LED bulb) & 24.60 & 26.60 & 34.80 & 45.60 \\ \hline
\end{tabular}
\end{description}

\begin{description}
\item[\quad Table K: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$E$ & 0 & 10 & 50 & 100 \\ \hline
$T$ (pay as-you-go) & 20 & 170& 770 & 1,520   \\ \hline
$T$ (member) & 150 & 350 & 650 & 1,150 \\ \hline
\end{tabular} 
\end{description}

\bigskip 

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Intercepts and direct proportionality -- Answers to exercises} % 4.3

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % different runners
\begin{enumerate}
\item $\approx 3.108$ miles
\item 7 min, 8 min, 3 min, 5 min, 0 min, 9 min
\item Galen, Aziz, Hitomi, Yannick, Olga, Fiona
\end{enumerate}

\item % snowstorm
\begin{enumerate}
\item $H=$ time (hours since 10 a.m.) $\sim$ indep \\ $S=$ total snow depth (inches) $\sim$ dep 
\item $S=4+\nicefrac{2}{3}H$ or $S=4+.6667H$
\item 4:00 a.m.
\item $T=$ time (hours since 4 a.m.) $\sim$ indep
\item $S=\nicefrac{2}{3}T$ or $S=.6667T$
\item $\checkmark$
\item $H$ starts at 10 a.m.\ so the intercept represents that there were 4 inches at 10 a.m.\  but $T$ starts at 4 a.m.\ so the intercept represents that there were 0 inches at 4 a.m. 
\end{enumerate}

\item % Paloma's beach
\begin{enumerate}
\item 435 feet
\item $T=$ time (from 60 years ago) $\sim$ indep \\ $D=$ depth of beach (feet) \\ $D=435 - 3.75T$
\item in another 8 years
\item Graph
\item Slope is -3.75 feet/year; it's how much beach depth is lost to erosion each year \\ Intercept is 435 feet; it's how deep the beach was 60 years ago.
\end{enumerate}

\item % Clyde's bricks
\begin{enumerate}
\item 17 pounds 
\item $B=$ number of bricks $\sim$ indep \\ $W=$ total weight (pounds) $\sim$ dep \\ $W=17 + 4.5B$
\item 152 pounds
\item 42 bricks
\item Graph, $\checkmark$
\end{enumerate}

\item % convenience pass bus
\begin{enumerate}
\item \$.6475/ride
\item No, it's just a discounted rate of 64.75\textcent~ per ride.
\item direct proportionality
\end{enumerate}

\item %baking cookies
\begin{enumerate}
\item 18 minutes
\item 2 hours and 18 minutes
\item $B=$ number of batches $\sim$ indep \\ $T=$ time to make cookies (min) $\sim$ dep \\ $M = 18 + 12B$
\item Slope is 12 minutes/batch; it's how long each batch bakes. \\ Intercept is 18 minutes; it's how long it takes to prepare the dough.
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Slopes -- Answers to exercises} % 4.4

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % paper by the case, revisited
\begin{enumerate}
\item \$42.50 per case
\item 55 pounds
\item delivered \$54.99 
\item $C=42.50N$
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$N$ &4 & 13\\ \hline
$C$ (delivered) &  174.96 & 534.87  \\ \hline
$C$ (store)& 170.00 & 552.50 \\ \hline
\end{tabular}
\item Graph
\item For 6 cases or more, delivered is cheaper.
\end{enumerate}

\item % garbage in U.S.
\begin{enumerate}
\item $\approx$ 3.61 million tons per year
\item $T=$ amount of garbage (million tons) $\sim$ dep \\ $Y=$ time (years since 1960 $\sim$ indep \\ $T=3.61Y + 88.1$
\item $\approx$ 268.6 million tons \\ $\approx$ 304.7 million tons
\item in 2019
\item There was more recycling and composting, or maybe because of the downturn in the economy?
\end{enumerate}

\item % Elmer's pension
\begin{enumerate}
\item 10\%
\item \$930.64
\item $P = 930.64-.10H$
\item $T = 93064 + .90H$
\item \$299.29
\item at least 37 hours
\end{enumerate}

\item % truck rental
 $D=$ distance (miles) $\sim$ indep \\ $R=$ cost of rental (\$) $\sim$ dep \\ $R=20 + .35D$ \\ \$23.50

\item % median household income
\begin{enumerate}
\item \$429/year
\item $Y=$ time (years since 2008) $\sim$ indep \\ $M=$ median household income (\$) $\sim$ dep \\ $M = 50,303-429Y$
\item in 2014
\item Graph, $\checkmark$
\end{enumerate}

\item % rising sea levels
\begin{enumerate}
\item $L=$ sea level (mm above historical level) $\sim$ dep \\ $Y=$ time (years since 2005 $\sim$ indep 
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$Y$ & 0 & 7\\ \hline
$L$ & 51.7 & 73.4  \\ \hline
\end{tabular}
\item 3.1 mm/year 
\item $L=51.7+3.1Y$
\item in 2015
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Fitting lines to data -- Answers to exercises} % 4.5

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % Thanh's road salt, revisited
\begin{enumerate}
\item  $\checkmark$
\item $\checkmark$
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$L$ & 600 & 1,000 \\ \hline
$S$ & 8,915 & 16,795  \\ \hline
\end{tabular}
$\checkmark$
\end{enumerate}

\item % wild rice
\begin{enumerate}
\item  Graph
\item $W = 4,296-65.17T$
\item 1,298 acres, 1,689 acres
\item Graph, seems too low and/or too flat
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$T$ & 35 & 40 & 47\\ \hline
$W$ & 2,188 & 1,776 & 1,199 \\ \hline
\end{tabular}

Graph
\item Down slightly to 1,281 acres. \\ 
Up quite a bit to 1,776 acres
\end{enumerate}

\item % garbage
\begin{enumerate}
\item Graph
\item Graph
\item Graph, later because it's not as steep a line.  That means it will take longer to increase to 300 million tons.
\end{enumerate}

\item % oil changes
\begin{enumerate}
\item Graph
\item Graph
\item $R=625-70N$ \\ \$625, \$65
\item 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$N$ & 0 & 4 & 8\\ \hline
$R$ & 733 & 352 & \cancel{-29}  \\ \hline
\end{tabular}

Graph
\item \$733, \$0 I guess (was negative)
\end{enumerate}

\item % sports movies
\begin{enumerate}
\item Graph, too flat (almost decreasing)
\item Graph, too high and maybe too steep
\item Graph, yes!
\item If a movie costs more, then maybe it has a better screenplay or actors or special effects or advertising and so it earn more.
\end{enumerate}

\item % meat and chicken
\begin{enumerate}
\item Graph
\item Beef line goes through 1985 and 2009 points. \\
Chicken line goes through 1975 and 1995 points.
\item in 2005?
\item in 2004, slightly earlier than estimated
\item Beef consumption has been fairly constant over the past 40 years.
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{A closer look at exponential equations} % Chapter 5

\bigskip

\subsection {Modeling with exponential equations -- Answers to exercises} % 5.1

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % sterling, revisited
\begin{enumerate}
\item around 1982
\item in 1982
\item in 1982
\end{enumerate}

\item % Mrs. Nystrom's social security
\begin{enumerate}
\item $S =$ Mrs.\ Nystrom's social security benefit (\$) $\sim$ dep \\ $Y=$ time (years since 2009) $\sim$ indep \\ $S=746.17 \ast 1.04^Y$
\item \$1,148.69
\item in 2014
\item in 2017
\end{enumerate}

\item % drawing app
\begin{enumerate}
\item \begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
$W$ & 0 & 2 & 4 & 6\\ \hline
$N$ & 2 & 5 & 12 & 30  \\ \hline
\end{tabular}
\item 8 weeks
\item 7.55 weeks
\item Graph
\end{enumerate}

\item % population Minnesota
\begin{enumerate}
\item 1.005
\item 5.33 million, 5.58 million
\item $M=5.2 \ast 1.005^Y$ \\ $Y =$ time (years since 2006) $\sim$ indep \\ $M=$ population of Minnesota (million) $\sim$ dep
\item See \textbf{Table L}, below.
\item Graph
\item in 2145 (if model continues that long, which is doubtful)
\end{enumerate}

\item % Um Archivo earnings
\begin{enumerate}
\item in 2015, in 2018
\end{enumerate}

\item % reality tv
growth factor is 1.072 \\ $V=$ number of viewers (millions of households) $\sim$ dep \\ $Y=$ time (years since 1990) $\sim$ indep \\ $V = 2.5 \ast 1.072 ^Y$ \\ around 5 million, around 10 million

\end{enumerate}
\end{multicols}

\bigskip

\begin{description}
\item[\quad Table L: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c|c |c|}\hline
year & 2006 & 2008 & 2010 & 2012 & 2014 & 2016 & 2018 & 2020 \\ \hline
$Y$ &  0 & 2 & 4 & 6 & 8 & 10 & 12 & 14 \\ \hline
$M$ & 5.2 & 5.25 & 5.30 & 5.36 & 5.41 & 5.47 & 5.5.2 & 5.58\\ \hline
\end{tabular} 
\end{description}

\bigskip 

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Exponential growth and decay -- Answers to exercises} % 5.2

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % caffeine, revisited (red bull)
\begin{enumerate}
\item $C=$ caffeine (mg) $\sim$ dep \\ $H=$ time (hours) $\sim$ indep \\ $C= 160 \ast .88^H$
\item about 5 \nicefrac{1}{2} hours
\item The caffeine would leave quicker, so it wouldn't take as long to get rid of half of it.  So, the half-life would be shorter.
\end{enumerate}

\item % pop bacteria
\begin{enumerate}
\item $B=$ number of bacteria $\sim$ dep \\ $D=$ time (days) $\sim$ indep 
\item See \textbf{Table M}, below.
\item $B=2,000 \ast 3^D$
\item 3
\item See \textbf{Table N}, below.
\item 5 days, 16 hours
\item Graph, see \textbf{Table M}, below.
\end{enumerate}

\item % Tenzin falling house prices
\begin{enumerate}
\item $Y=$ time (years since he bought house $\sim$ indep \\ $H=$ value of Tenzin's house (\$) $\sim$ dep \\ $H=291,900 \ast .959^Y$
\item 5 years 
\item 4 years, 8 months
\end{enumerate}

\item % wetlands /contaminants
\begin{enumerate}
\item 2 grams, .5 grams, .125 grams
\item $M=$ time (months) $\sim$ indep \\ $C=$ amount of contaminant (grams) $\sim$ dep \\ $C = 8 \ast .25^M$
\item .25
\item 8 months, 1 year
\item Graph
\end{enumerate}

\item % Hibbing population
\begin{enumerate}
\item .996
\item $H=$ population Hibbing (people) $\sim$ dep \\ $Y=$ time (years since 2000) $\sim$ indep \\ $H = 17,071 \ast .996^Y$
\item 16,400 
\item faster than expected, so more than .4\% decay rate
\end{enumerate}

\item % donations food shelf
\begin{enumerate}
\item $T=$ time (years since last year) $\sim$ indep \\ $L=$ donations to local food shelf (pounds) $\sim$ dep 
\item $L=3,400 \ast 1.35^T$
\item $\approx$ 4,600 pounds, $\approx$ 6,200 pounds, $\approx$ 8,400 pounds
\item 2 years, 4 months
\item in 4 more years
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip

\begin{description}
\item[\quad Table M: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c|c |c|}\hline
$D$& 0 & 1 & 2 & 3  & 4 & 5 & 6\\ \hline
$B$& 2,000 & 6,000 & 18,000 & 54,000  & 162,000 & 486,000 & \cancel{1.458,000} \\ \hline
\end{tabular} 
\end{description}

\begin{description}
\item[\quad Table N: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c|c |c|}\hline
$D$  & 10 & 20 & 30\\ \hline
$B$& $1.2\times 10^{8}$ & $6.9 \times 10^{12}$ & $4.1 \times 10^{17}$ \\ \hline
\end{tabular} 
\end{description}

\bigskip 

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Growth factors -- Answers to exercises} % 5.3

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % childhood obesity, revisited
\begin{enumerate}
\item 1.0412, 4.12\%
\item $C=1.1 \ast 1.0412^Y$
\item around 3.4 out of every ten children
\item Graph
\end{enumerate}

\item 
\begin{enumerate}
\item % thick glass
25\% decrease per inch
\item % pop bacteria
200\% increase per day
\item % wetlands /contaminants
75\% decrease per month 
\end{enumerate}

\item % Whitney's antique table
\begin{enumerate}
\item 1.06496121\ldots
\item $V=$ value of Whitney's table (\$) $\sim$ dep \\ $A=$ age of table (years) $\sim$ indep \\ $V=560 \ast 1.065^A$ \\ $\approx$ \$6,900
\end{enumerate}

\item % morphine
\begin{enumerate}
\item 10 mg
\item 3 days
\item .46415888\ldots
\item $\approx$ 53.6\% 
\item $M=$ amount of morphine (mg) $\sim$ dep \\ $H=$ time (hours) $\sim$ indep \\ $M = 100 \ast .4641588^H$ $\checkmark$
\item just under 22 hours
\item Graph
\end{enumerate}

\item % unemployment
\begin{enumerate}
\item 1.0665051\ldots
\item $M=$ time (months) $\sim$ indep \\ $U=$ number of unemployed adults $\sim$ dep \\ $U = 20,517 \ast 1.0665051^M$
\item 57,481 unemployed
\item See \textbf{Table O}, below.  Graph.
\end{enumerate}

\item % MN wetlands decrease
\begin{enumerate}
\item $Y=$ time (years since 1850) $\sim$ indep \\ $A=$ wetlands (acres) $\sim$ dep \\ .99547987\ldots \\ $A=18.6 \ast .99547987^Y$
\item $T=$ time (years since 2003) $\sim$ indep \\ $A=$ wetlands (acres) $\sim$ dep \\ 1.0146439\ldots \\ $A=9.3 \ast 1.0146439^T$
\end{enumerate}

\end{enumerate}
\end{multicols}

\bigskip

\begin{description}
\item[\quad Table O: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
time & 10 mo ago & now & in 6 mos & in 1 year & in 2 yrs \\ \hline
$M$ & 0 & 10 & 16 & 22 & 32 \\ \hline
$U$ & 20,517 & 39,061 & 57,481 & 84,586 & 161,038  \\ \hline
\end{tabular}
\end{description}

\bigskip

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Linear vs.\ exponential models -- Answers to exercises} % 5.4

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % weight loss is like car depreciation?
$M=$ time on diet (months) $\sim$ indep \\ $W=$ weight (pounds) $\sim$ dep
\begin{enumerate}
\item \textbf{linear:} $W=165-4M$ \\ $\approx$ 12 months if linear \\
\textbf{exp:} $W=165 \ast .97514491^M$ \\ $\approx$ 13 months if exponential
\item \textbf{linear:} $W =253-6M $ \\$\approx$ 13 months if linear \\
\textbf{exp:} $W = 253\ast .97569882^M$ \\ $\approx$ 15 months if exponential
\item You tend to lose more weight at first and then it slows down.
\end{enumerate}

\item %CFC 
\begin{enumerate}
\item 
\begin{enumerate}
\item 8.295 ppt/year
\item $C=11.1 + 8.295Y$
\item $\approx$ 467 ppt
\item linear
\end{enumerate}
\item 
\begin{enumerate}
\item 1.14850400\ldots
\item $C=11.1\ast1.14850400^Y$
\item $\approx$ 22,500 ppt
\item exponential
\end{enumerate}
\end{enumerate}

\item % recommended caloric intake 
\begin{enumerate}
\item $C=$ recommended intake (calories) $\sim$ dep \\ $A=$ age (years since 30) $\sim$ indep 
\item $\approx$ 1,960 calories, $\approx$ 1,780 calories, $\approx$ 1,600 calories
\item $\approx$ 1,962 calories, $\approx$ 1,782 calories, $\approx$ 1,618 calories
\item The two models give almost exactly the same numbers.
\end{enumerate}

\item % price movie ticket
\begin{enumerate}
\item $T=4.35+.22375Y$ \\ \$8.83, \$11.06
\item $T=4.35\ast1.038243^Y$ \\ \$9.21, \$13.41
\item Graph
\item linear
\end{enumerate}

\item % asthma
In 2015, linear is 189 million and exponential is 227 million. \\
In 2030, linear is 251 million and exponential is 411 million. \\
Graph

\item % hybrid car sales
\begin{enumerate}
\item $H=17+11,501Y$ \\ $\approx$ 138,000 cars sold
\item $H = 17 \ast 12.663239^Y$ \\ $\approx$ 290 trillion cars, which is not realistic!
\item Sales might be down because of the economy, or because many people who want to own a hybrid already do.
\end{enumerate}

\end{enumerate}
\end{multicols}

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}

\subsection {Logistic and other growth models -- Answers to exercises} % 5.5

\begin{multicols} {2}
\begin{enumerate}
\setcounter{enumi}{4}

\item % revisit flu in dorms
$\checkmark$

\item % calls about plastic -- logistic
\begin{enumerate}
\item See \textbf{Table P}, below.
\item January
\item Graph
\item People stopped calling.
\end{enumerate}

\item % calls about plastic -- saturation
\begin{enumerate}
\item See \textbf{Table Q}, below.
\item 98\%, 99.9\%
\item day 69 (March 11)
\item day 69 (March 11)
\end{enumerate}

\item % infant height
\begin{enumerate}
\item 3.5 kg, 4.3 kg, 6.3 kg, 10.1 kg
\item 7 lb \& 11 oz, 9 lb \& 7 oz, 13 lb \& 15 oz, 22 lb \& 2 oz
\item Graph. \\ The curve levels off which would mean that the infant does not gain any more weight.
\end{enumerate}

\item% Rodney's fish
\begin{enumerate}
\item See \textbf{Table R}, below.
\item Graph
\item saturation
\item logistic
\end{enumerate}



\end{enumerate}
\end{multicols}

\bigskip

\begin{description}
\item[\quad Table P: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
day & Jan 1 & Feb 1 & Mar 1 & Apr 1 & May 1 & Nov 8 \\ \hline
$D$ & 0 & 31 & 59 & 90 & 120 & 311\\ \hline
$C$ & 7 & 741 & 837 & 837 & 837 & 837  \\ \hline
\end{tabular}
\end{description}

\begin{description}
\item[\quad Table Q: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
day & Jan 1 & Mar 1 & Aug 1 & Oct 1 & Nov 8\\ \hline
$D$ & 0 & 59 & 212 & 273 & 311\\ \hline
$P$ &  13 & 46 & 84 & 90 & 93 \\ \hline
\end{tabular}
\end{description}

\begin{description}
\item[\quad Table R: \quad] 
\begin{tabular} {|c| |c|c |c|c |c|c |c|}\hline
year & 10 yrs ago & now & in 10 yrs & in 20 yrs & in  30 yrs \\ \hline
$Y$ & 0 & 10 & 20 & 30 & 40 \\ \hline
$B$ (saturation) & 400 & 3,700 & 3,975 & 3,998 & 4,000 \\ \hline
$B$ (logistic) & 400 & 2,285 & 3,765 & 3,979 & 3,998 \\ \hline
\end{tabular}
\end{description}

\bigskip

\begin{center}
\line(1,0){300} %\line(1,0){250}
\end{center}
