
\setcounter{chapter}{-1}
\chapter{Prelude} 

You have been learning and using mathematics your entire life.  And not just in school.  At home, on the job, and as you go about your day. Perhaps you were pretty good at math in school, or perhaps it was a huge struggle for you, or something in between.  You might be upset that you have to take another math class or you might be optimistic that this time around it's gonna make sense -- I'm hoping it will!

While you have studied mathematics for many years in school, I suspect there's a lot of mathematics you haven't seen or used in awhile. That's completely normal.  The good news is you don't need to know or remember \emph{everything} you've ever seen in math classes to learn algebra.  We will review what we need when we need it.

These preludes present the key ideas about arithmetic operations (adding, subtracting, multiplying and dividing) using a calculator and the order of operations (PEMDAS anyone?), decimal numbers, rounding, percentages, fractions, powers, and roots (like square roots).  Think of the preludes as bumping into an old friend.  You might share a few niceties, but you'll need meet up for lunch another day to catch up more.  So, while these preludes present the basics, we will revisit and use these topics throughout the course in more detail and you'll have more chance to practice then.

It's important that you know that everyone can learn mathematics.  Really, everyone.  That means you.  I hope this is a wonderful and successful experience for you and that by the end of this course you recognize your new and improved mathematical superpowers.  I believe you can do it.


